Coupled response of stomatal and mesophyll conductance to light enhances photosynthesis of shade leaves under sunflecks

### 2016-09-11 PCE_paper/review_response.Rmd

---
output:
  word_document:
    reference_docx: manuscriptstyle_pce2.docx
  pdf_document: default
  html_document:
    number_sections: yes
---

@@

```{r, results='asis', echo=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

Keith A. Mott
\
Editor-in-Chief
\
Plant, Cell and Environment

\
Dear Professor Mott,

We are grateful for the opportunity to resubmit our manuscript. We thank the reviewers for providing insightful comments to improve our manuscript. We have responded in detail to each comment, which we feel has improved the overall quality of the manuscript. Below, we have itemized our response to each comment from the Editor and both referees.

### **Editor**
\
- *Editor: 1. Were these mature leaves and therefore isobilateral? Does this perhaps affect the outcome? The fact that LMA did not vary with shade is different from many other reports for dorsiventral leaves.*

@@

\
*Response*: The leaves studied in this experiment were very likely juvenile leaves, as the *E. tereticornis* trees were grown for ~ 1 year. Consequently, the fact that LMA was similar within the canopy is different from other reports as the editor notes. Similar LMA within canopies has been noted before in Eucalypt plantations (Nouvellon et al. 2010), which is cited in the discussion. We have expanded the discussion to address the possible fixed controls of LMA on leaf physiology (see reply to Referee # 2 specific comments for full details).

\
- *Editor: 2. (Stylistic) I found the colors to be distracting. Judicial use of color can help clarify data but unnecessary use of color can be distracting.*

\
*Response*: We feel that the use of color was necessary to visualize the three leaf types (sun, shade, shade+light), while also differentiating between the two temperature treatments. We attempted to use only symbols, which makes it near impossible to distinguish all of the treatments. It also complicates visualization of confidence intervals on scatter plots. The choice of color palette is also used to properly colors for color-blind readers.

@@

\
*Editor: 3. The use of Jmax is consistent with the original use but not consistent with more modern use of the J that is found when light is extrapolated to infinity (similar to the concept of Vmax. Buckley and Diaz-Espejo recently published on this and given the use of Jmax in global models it might be good to limit the use of Jmax to the extrapolated number meaning that the J from an ACi curve analysis should be called something else (Jhigh? Jsat?) not sure specifically what to recommend, I avoid the issue by using just J but am not sure that is the best solution either.*

\
and

\
*4. I don't like expressing Vcmax on a "Ci" basis. This again changes the meaning of Vcmax from the extrapolated maximum rubisco rate (biochemically correct) to enzyme and biophysical diffusion components (no longer biochemically a Vmax and making two different definitions of Vcmax. Given that mesophyll conductance is a central part of this manuscript I don't understand why some reasonable constraints cannot be put on the A-Ci analysis so that parameters can be estimated based on Cc. I can accept that mesophyll conductance derived from ACi curves may not be sufficiently reliable, but the measurements made with the TDL should allow some reasonable Gm's to be applied to the ACi curves to avoid a Vcmax that is both biochemistry and biophysics.*

\
*Response*: After careful thought we have decided to remove parameters derived from AC~i~ curves from the manuscript. Our main goal is to show that sun and shade leaves differed in their photosynthetic capacity, regardless of whether these differences are due to biochemistry or leaf structure (which we cannot satisfactorily address with the available data). As such, we have decided to instead present photosynthetic rates at saturating CO~2~ concentrations and light (*A*~max~) for the different leaf types. This provides a simple metric of photosynthetic capacity, and shows that shade leaves have a lower photosynthetic capacity. We have adjusted the Methods (which is now shorter), as well as Figure 2 which now shows the relationship between *A*~max~ and leaf *N*~a~. *A*~max~ values are also reported in Table 1.

@@

\
We acknowledge that AC~i~ response curves were initially conceived with the notion that *g*~m~ was large enough to be ignored, which we now know is untrue. We do think there are some practical concerns when fitting A-Cc curves, and we hope this data set can eventually help to address this issue, but it is beyond the scope of this paper.

\
*Editor: 5. Again because the leaves are isobilateral, to what degree is the increase in gm in high light a reflection of photosynthesis happening over a greater depth of leaf?*

\
*Response*: This question raises theoretical issues about how/if photosynthetic machinery are optimally distributed along intraleaf light gradients. To our knowledge, these issues have yet to be resolved empirically and are still a relevant question for most leaf-level experiments. For example, it is possible that *J*~max~ and/or *V*~cmax~ could be partitioned within the leaf according to light gradients (see Lloyd et al. 1992), which may affect estimates of *g*~m~ at low light.  Alternatively, greater light penetration into the leaf but longer path lengths for CO~2~ diffusion might interact to affect estimates of *g*~m~ at high light. Although we acknowledge that this question is relevant to this study and many others, we do not have the tools to address it further in the manuscript.

@@

### **Referee #1**

\
*Reviewer: Campany et al have presented a work on spatial distribution of leaf gas exchange variables and parameters as a function of light regime and how they reflect an optimal distribution of resources in the canopy. The most remarkable result of this study is the occurrence of a higher gs in shaded leaves than in sunny ones.  It is not clear to me the advantages of shaded leaves in having higher gs than ones at sun. I can see the point of enhancing An when sunflecks, but if the whole day is integrated it would result in a dramatic reduction of WUE, since shaded leaves with stomata open will transpire more or similarly to sunny ones (assuming that sunny leaves have higher temperature and therefore higher leaf-to-air VPD. Honestly, I cannot see any optimisation for canopy C in here, if this would be the case. As authors discussed, perhaps we should considered a dynamic light environment to understand the advantages of this strategy, but still I find difficult to conciliate with an optimal distribution of resources, in most cases very limiting. Data shows that there was an increase of gs, and hence, water use, without an enhancement of KL, and the relationship between gsmax and Vcmax, which usually found in canopies, but in this study is broken was broken.*

\
*Response*: Due to the concern over cuvette measured transpiration, water potentials and subsequent *K*~L~ calculations, we have re-examined the calculation and subsequent analyses of *K*~L~ in the manuscript. Initially, *K*~L~ was calculated with transpiration rates of shade leaves under low light conditions. As shade leave were allowed to stabilize under high light conditions before they were detached and measured for water potential, the leaf water status reflected these conditions and not those of low light. We have corrected the calculation of *K*~L~ and made changes in the results and discussion sections. We have also clarified the calculation in the Methods section.

\
After the recalculation, *K*~L~ is now significantly higher in shade leaves compared to sun leaves, which makes sense to maintain hydraulic supply to the observed stomatal behavior. We believe this actually makes the optimization of resources story much clearer throughout the manuscript, and we greatly appreciate the referee's comments. These corrected values also allow us to address the possible coordination of *K*~L~ and *g*~m~ reported by Flexas et al. 2013.

@@

\
*Reviewer:  Since the theory of considering a dynamic light environment is very appealing and novel, I would suggest to the authors to make diurnal cycles of light, gs and An at different locations in the canopy and check this hypothesis. They have all the equipment necessary to do so and they could integrate a whole day and see if it makes sense.*

\
*Response*: The trees in the whole-tree chambers were harvested in May 2014, shortly after our measurements were complete. Therefore, the suggestion to make diurnal measurements is no longer possible. We do agree with the reviewer that such detailed measurements would make for an interesting follow-up study.

\
*Reviewer: Another potential flaw I can see is the way KL was measured in this study. I would say it is not acceptable if we want to reach conclusions about the spatial distribution of K within the canopy. The assumed steady-state in the IRGAs chamber while measuring transpiration might not be the same one of the leaves in which leaf water potential was measured. Likely it is not, since leaves modify the VPD in the chamber respect to the one under natural conditions. Additionally, it seems that KL was not estimated from measurements of leaf water potential and E taken on the same leaf. At least, it is not specified. This can be subjected to errors in the estimation of KL. Finally, the gradient of water potential was considered between the soil (assuming leaf water potential as a good surrogate of matric potential) and the leaf. So authors were measuring the plant hydraulic conductance, from leaf to root. It is normal that it is similar in all leaves of the plant.*

\
*Response*: The main issue in this comment is whether or not leaf gas exchange and water potential were measured on the same leaves. As explained in the section "Leaf Nitrogen and hydraulic conductance", they were indeed measured on the same leaves. Leaves used for gas exchange were indeed sampled and then immediately measured for midday water potential. The method we used to estimate KL is standard, and presented as a commonly used method in Flexas et al. 2013, which compared the coordination of leaf hydraulic conductance and mesophyll conductance. We now cite this work in the Methods section, as justification for the methods used.

@@

\
*Response*: We thank the reviewer for the positive comments, and agree that the points raised were the most surprising results in the study.

\
*Reviewer: I have two significant concerns about it. First, the mesophyll conductance calculation apparently assumed that the respiratory and photorespiratory carbon pools were at isotopic steady-state. This assumption is most problematic in the leaves moved from shade to full sunlight, where the cc/ca likely fell substantially. It is not clear that the carbon pools supporting the respiratory and photorespiratory flux terms would have been able to shift to the new photosynthate in the brief acclimation period (ca. 25 min) of these experiments. This issue was raised by Wingate et al. in a nice PCE ms. several years ago (WINGATE et al., 2007). Perhaps the modeling could be conducted with the assumption that old photosynthate makes up these pools and a sensitivity analysis conducted to determine whether it matters. Alternatively, perhaps the TDL data can be used to describe the shift in isotopic composition of the net flux; an asymptote might be used as evidence of a new steady-state having been reached. It would be great for the community if you were able to use a description of this shift to estimate the turnover time of the pools!*

\
*Response*: The reviewer is correct that we did not account for 'old' carbon in respiratory pools. We have now recalculated the respiratory fractional component 'e' (according to Evans and von Caemmerer 2013), which addresses this issue (which was described by Wingate et al. 2007). The correction increased values of *g*~m~ but does not alter observed patterns. The correct calculation of 'e' has been added to the supplementary methods.  This correction also addresses the next comment.

\
We agree with the reviewer that "Perhaps the modeling could be conducted with the assumption that old photosynthate makes up these pools and a sensitivity analysis conducted to determine whether it matters". However we think this idea is well outside the scope of the current study, but it is an issue that could be addressed in future work.

### 2016-09-11 PCE_paper/manuscript_gmes.Rmd
manuscript edits

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Set Flag to include figures in text or not.
includeFigs <- FALSE
```

@@

source("gm_pce/manuscript_preload.R")

# Make figure PDFs.
source("gm_pce/make_figures.R")

@@

# Abstract
Light gradients within tree canopies play a major role in the distribution of plant resources that define the photosynthetic capacity of sun and shade leaves. However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting canopy carbon and water exchange. To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole-tree chambers. Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*A*~max~, i.e. *A*~n~ at saturating [CO~2~]), but *g*~s~ was similar. When light intensity was temporarily increased for shade leaves to match that of sun leaves, both *g*~s~ and *g*~m~ increased, and *A*~n~ increased to values greater than sun leaves. Here we show that dynamic physiological responses of shade leaves to altered light environments have implications for up-scaling leaf level measurements and predicting whole canopy carbon gain. Despite exhibiting reduced photosynthetic capacity, the rapid up-regulation of *g*~m~ with increased light enables shade leaves to respond quickly to sunflecks.

@@

\
In sun leaves, *A*~n~ and *g*~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory (Fig. 5). However, lower rates of *A*~n~ in shade leaves were not coupled with decreases in *g*~s~, leading to the observed decreases in ITE. This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [e.g. @prentice2014balancing; @lin2015optimal]. Consistent with our results, Buckley et al. [-@2014] found that shade leaves exhibited higher *g*~s~ than expected based on their environment. In our study, higher *g*~s~ in shade leaves may in part be explained by higher *K*~L~, which effectively reduces the cost of water transport, allowing more profligate water use in shade leaves.

@@

```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, eval=includeFigs}
 source("master_scripts/aci_vcmaxnitro.R")

>>>>

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Set Flag to include figures in text or not.
includeFigs <- TRUE
```

@@

source("PCE_paper/manuscript_preload.R")

# Make figure PDFs.
source("PCE_paper/make_figures.R")

@@

# Abstract
Light gradients within tree canopies play a major role in the distribution of plant resources that define the photosynthetic capacity of sun and shade leaves. However, biochemical and diffusional constraints on gas exchange in sun and shade leaves in response to light remain poorly quantified, but critical for predicting canopy carbon and water exchange. To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with concurrent measurements of carbon isotope discrimination to measure net leaf photosynthesis (*A*~n~), stomatal conductance (*g*~s~) and mesophyll conductance (*g*~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole-tree chambers. Compared to sun leaves, shade leaves had lower *A*~n~, *g*~m~, leaf nitrogen and photosynthetic capacity (*A*~max~, i.e. *A*~n~ at saturating [CO~2~]), but *g*~s~ was similar. When light intensity was temporarily increased for shade leaves to match that of sun leaves, both *g*~s~ and *g*~m~ increased, and *A*~n~ increased to values greater than sun leaves. We show that dynamic physiological responses of shade leaves to altered light environments have implications for up-scaling leaf level measurements and predicting whole canopy carbon gain. Despite exhibiting reduced photosynthetic capacity, the rapid up-regulation of *g*~m~ with increased light enables shade leaves to respond quickly to sunflecks.

@@

\
In sun leaves, *A*~n~ and *g*~s~ were strongly correlated, exhibiting behaviour broadly consistent with optimal stomatal theory (Fig. 5). However, lower rates of *A*~n~ in shade leaves were not coupled with decreases in *g*~s~, leading to the observed decreases in ITE. This is significant as optimal stomatal regulation to balance C gain with water loss has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [e.g. @prentice2014balancing; @lin2015optimal]. Consistent with our results, Buckley et al. [-@buckley2014] found that shade leaves exhibited higher *g*~s~ than expected based on their environment. In our study, higher *g*~s~ in shade leaves may in part be explained by higher *K*~L~, which effectively reduces the cost of water transport, allowing more profligate water use in shade leaves.

@@

```{r amaxnitro, fig.width=6, fig.height=6, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, eval=includeFigs}
 source("master_scripts/amax_nitro.R")
